---
title: "ANA 515 Assignment 2"
author: "Grace Zeng"
date: "6/16/2024"
output:
  html_document: null
  theme: journal
  pdf_document: default
---

### Description of the data：
The dataset* selected for this analysis comes from the FBI and the Southern Poverty Law Center (SPLC) and the research question is: What are the trends in hate crime rates by state, and what state-level factors are associated with these variations?

Data collection: The federal government does not systematically track hate crimes, as reporting to the FBI is voluntary. The Southern Poverty Law Center relies on media accounts and self-reports to evaluate the situation. Both sources provide publicly available data.

*The dataset is found on [GitHub](https://github.com/fivethirtyeight/data/tree/master/hate-crimes)
 and it is in csv. format.  More background information can be found in this [article](https://fivethirtyeight.com/features/higher-rates-of-hate-crimes-are-tied-to-income-inequality/).


---


```{r setup, include= FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("tidyverse")
#install.packages("knitr")
#install.packages("bslib")

library(tidyverse)
library(knitr)
library(bslib)
```

```{r GitHubdata, include= FALSE}
#task 2: Loading the data：
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/hate-crimes/hate_crimes.csv"
hate_crimes <- read_csv(url)
glimpse(hate_crimes)
# I used read_csv function and this function comes from the tidyverse package
```

### Get a glimpse of the data
```{r glimpsedata, echo = FALSE}
#task 2: Loading the data：
glimpse(hate_crimes) 
```

---

```{r datacleaning, include= FALSE}
#task3: Clean the data
#arrange
hate_crimes_arrange_by_income <- arrange(hate_crimes, median_household_income)

#filter
morethan50_share_voters_voted_trump <-filter(hate_crimes, share_voters_voted_trump>0.5)

```


### Characteristics of the data

This dataframe has has `r nrow(hate_crimes_arrange_by_income)`  rows and `r ncol(hate_crimes_arrange_by_income)` columns. The names of the columns and a brief description of each are in the table below:

```{r tabledescription, echo = FALSE}
#task 4 Characteristics of the data
library(knitr)

# Your dataset
data_subset <- hate_crimes_arrange_by_income

# Description of columns
columns_summary <- data.frame(
  Columns = colnames(data_subset),
  Description = c(
    "State name", 
    "Median household income, 2016",
    "Share of the population that is unemployed (seasonally adjusted), Sept. 2016",
    "Share of the population that lives in metropolitan areas, 2015",
    "Share of adults 25 and older with a high-school degree, 2009",
    "Share of the population that are not U.S. citizens, 2015",
    "Share of white residents who are living in poverty, 2015",
    "Gini Index, 2015",
    "Share of the population that is not white, 2015",
    "Share of 2016 U.S. presidential voters who voted for Donald Trump",
    "Hate crimes per 100,000 population, Southern Poverty Law Center, Nov. 9-18, 2016",
    "Average annual hate crimes per 100,000 population, FBI, 2010-2015"
  )
)

# Create a table with descriptions
kable(columns_summary, caption = "Description of Columns in hate_crimes_arrange_by_income Dataset")

```

### Summary of data subset

```{r subset, include= FALSE}
#task 5 Summary statistics
#subset of the dataset

subset_hate_crimes_arrange_by_income <- select(hate_crimes_arrange_by_income, state, median_household_income, share_unemployed_seasonal, avg_hatecrimes_per_100k_fbi)

```

Subset of the data: 
```{r subset_of_data,echo = FALSE }
head(subset_hate_crimes_arrange_by_income)
```

Summary of the subset:
```{r summary, echo = FALSE}
Summarytable <-summary(subset_hate_crimes_arrange_by_income) #creates the summary
Summarytable #prints the summary in your output
```

